A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-07 manuscript.Rmd
cite tidyverse

Modern data analysis is only feasible because researchers rely on much external software. <!-- AP: can we give examples of cool applications in psychology -->
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility. The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance . At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Containers directly address such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself. The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such. <!--AP can we add an example here with the non-reproducibility of an eigen problem solver - I think we discussed at some point in time when WOlfgang Viechtbauer sent us a link to a paper -->.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
There are two major strategies helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images for R based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they additionally need on top of these pre-made images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse <!-- AP add citations to all packages, please--> (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`. This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image. However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`

@@

```{r, include=FALSE}
knitr::write_bib(c(.packages(), "gert", "bookdown", "lavaan", "knitr", "targets", "renv"),
                 "packages.bib")
```

>>>>

Modern data analysis is only feasible because researchers rely on much external software. <!-- AP: can we give examples of cool applications in psychology -->
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility. The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance . At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Containers directly address such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself. The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such. <!--AP can we add an example here with the non-reproducibility of an eigen problem solver - I think we discussed at some point in time when WOlfgang Viechtbauer sent us a link to a paper -->.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
There are two major strategies helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images for R based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they additionally need on top of these pre-made images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`. This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image. However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`

@@

```{r, include=FALSE}
knitr::write_bib(c(.packages(), "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
                 "packages.bib")
```
